---
title: "Benthic Clustering & Ordination"
author: "Julia Briand"
date: "1/26/2022"
output: html_document
---
This script was used to prepare data for Spatio-temporal patterns in coral reef composition and function across an altered environmental gradient. Contact: Julia Briand (julia.briand@mail.mcgill.ca)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ncdf4)
library(fields)
library(tidyverse)
library(sf)
library(raster)
library(rasterVis)
library(rgdal)
library(biogeo)
library(vegan)
library(gclus)
library(ape)
library(labdsv)
library(rdaTest)
library(ggvegan)
library(nlme)
library(directlabels)
library(ggrepel)
library(maptools)
library(mvpart)
library(MVPARTwrap)
library(plyr)
library(MASS)
```

## Read in environmental and benthic data

```{r}
env <- read.csv("Env_Supp.csv", row.names = 2) 
spe <- read.csv("Species.csv", row.names = 2)
```

## Transform species (benthic cover) data

```{r}
spe.log <- log1p(spe)
spe.log.bray <- vegdist(spe.bray)
spe.mean.log <- log1p(spe.mean)
```

## Standardize environmental Data

```{r}
env.z <- decostand(env, method="standardize")
# the data are now centered (means~0)
apply(env.z, 2, mean) 
# the data are now scaled (standard deviations=1)
apply(env.z, 2, sd)
```

## K means clustering

```{r}
# PCoA Site Scores
mean_pcoa <- capscale(spe.mean.log ~ 1, dist="bray", add = TRUE, metaMDS = TRUE)
mean_scores <- scores(mean_pcoa)
site.scores <- mean_scores$sites

# K means
spe.kmeans <- kmeans(site.scores, centers = 2, nstart = 100000)
spe_grps <- as.data.frame(unlist(spe.kmeans$cluster))
names(spe_grps) <- (c("Group"))
spe_grps <- rownames_to_column(spe_grps, var = "Site")
spe_grps # clusters
```

## Multivariate Regression Trees

```{r}
benth.mrt <- mvpart(as.matrix(spe.log) ~SST.Mean + SPM + CHL + Fetch + Depth,env.z,
                    legend=FALSE, margin=0.01, cp=0, xv="pick",
                    xval=nrow(spe.log), xvmult=100, which=4, size = 5)

summary(benth.mrt)
printcp(benth.mrt)

benth.mvpart <- factor(benth.mrt$where)
levels(benth.mvpart) <- 1:length(levels(benth.mvpart))
benth.mvpart

rn <- rownames(spe.log)
rn <- as.data.frame(rn)
df <- cbind(rn, benth.mvpart)

write.csv(df, "temp_benth_mrt_grps.csv")
```

## Distance Based RDA

```{r}
# RDA
SST_rda <- capscale(spe.log ~ SST.Mean + SPM + CHL + Fetch + Depth,
                     env.z, add = TRUE, dist="bray")
summary(SST_rda, display=NULL)

# R2
SST_R2adj <- RsquareAdj(SST_rda)$adj.r.squared

# Model selection
ordiR2step(capscale(spe.log ~ 1, data=env.z, add = TRUE, dist = "bray"), scope= formula(SST_rda), direction= "forward", R2scope=TRUE, pstep=1000)


# Significance of Axes
anova.cca(SST_rda, step=1000)
anova.cca(SST_rda, step=1000, by="axis") 

# Site and species scores
SST_scores <- scores(SST_rda)

# Top 10 % of species scores along the first axis
abs_DB_SST <- as.data.frame(abs(SST_scores$species), keep.rownames=TRUE)
SST_DB_MDS1_spe <- abs_DB_SST[abs_DB_SST$CAP1 > quantile(abs_DB_SST$CAP1,prob=1-10/100),]
SST_DB_spe <- rownames(SST_DB_MDS1_spe)
SST_DB_spe
```

## PCoA

```{r}
benth_pcoa <- capscale(spe.log ~ 1, dist="bray", add = TRUE, metaMDS = TRUE)
benth_scores <- scores(benth_pcoa)

# Top 15 % of species scores along the first axis
abs_benth <- as.data.frame(abs(benth_scores$species), keep.rownames=TRUE)
benth_MDS1_spe <- abs_benth[abs_benth$MDS1 > quantile(abs_benth$MDS1,prob=1-15/100),]

# Top 15 % of species scores along the second axis
benth_MDS2_spe <- abs_benth[abs_benth$MDS2 > quantile(abs_benth$MDS2,prob=1-15/100),]
benth_spe <- rownames(benth_MDS1_spe)

pcoa_scores_dat <- as.data.frame(benth_scores$species)
pcoa_scores_dat[row.names(pcoa_scores_dat)[row.names(pcoa_scores_dat) %in% 
                                             row.names(benth_MDS1_spe)],]

# Plot top 15 % on first axis
pcoa_spe_dat <- as.data.frame(benth_scores$species[benth_spe,])

ggplot(pcoa_spe_dat, aes(MDS1, MDS2)) + 
  geom_segment(aes(x = rep(0,7),  y = rep(0,7), xend = MDS1, yend = MDS2), 
               data = pcoa_spe_dat, arrow=arrow(length = unit(0.3,"cm"))) +
  geom_text_repel(aes(label = rownames(pcoa_spe_dat)), 
                  data = pcoa_spe_dat, box.padding = 0.2, 
                  nudge_y = -0.01, nudge_x = -0.04) + 
  ggthemes::theme_base() +
  xlim(-2,2) + ylim(-2,2)

ggsave("benthic_scores.pdf", height=5, width=6)
```

## Set up tibbles

```{r}
env.tib <- as_tibble(env, rownames = "Site")
spe.tib <- as_tibble(spe, rownames = "Site")
```

## Temporal RDAs

```{r}
# Exposed
spe_exp <- spe.tib %>% filter(Year >= 2008) %>% select(-c(Year)) %>%  
  filter(Site %in% c("zapoeste", "isbasti", "tobobe", "bastinor")) %>% column_to_rownames("Site.Year") %>% select(-Site)

# Protected
spe_pro <- spe.tib %>% filter(Year >= 2008) %>% select(-c(Year)) %>%  
  filter(Site %in% c("caroldan", "Penivali", "iscaagua", "iscolon", "iscristo", "issobut", "Gallialm")) %>% column_to_rownames("Site.Year") %>% select(-Site)

# Subset Env Data

# Exp Env
env_exp <- env.tib %>% filter(Year >= 2008) %>% 
  filter(Site %in% c("zapoeste", "isbasti", "tobobe", "bastinor")) %>% column_to_rownames("Site.Year")

# Pro Env
env_pro <- env.tib %>% filter(Year >= 2008) %>%
  filter(Site %in% c("caroldan", "Penivali", "iscaagua", "iscolon", "iscristo", "issobut", "Gallialm")) %>% column_to_rownames("Site.Year")

# Partial dbRDA Exposed
exp_time <- capscale(log1p(spe_exp) ~ as.numeric(Year) + Condition(Site), env_exp,
                     add = TRUE, dist="bray")
summary(exp_time)
exp_R2adj <- RsquareAdj(exp_time)$adj.r.squared

# Partial DB RDA scores
exp_scores <- scores(exp_time)

# Top 15 % of species scores along the first axis (Year)
abs_exp <- as.data.frame(abs(exp_scores$species), keep.rownames=TRUE)
exp_RDA_spe <- abs_exp[abs_exp$CAP1 > quantile(abs_exp$CAP1,prob=1-15/100),]
exp_spe <- rownames(exp_RDA_spe)
exp_spe

# Partial dbRDA Protected
pro_time <- capscale(log1p(spe_pro) ~ as.numeric(Year) + Condition(Site), env_pro,
                     add = TRUE, dist="bray")
summary(pro_time)
pro_R2adj <- RsquareAdj(pro_time)$adj.r.squared

# Partial DB RDA scores
pro_scores <- scores(pro_time)

# Top 15 % of species scores along the first axis (Year)
abs_pro <- as.data.frame(abs(pro_scores$species), keep.rownames=TRUE)
pro_RDA_spe <- abs_pro[abs_pro$CAP1 > quantile(abs_pro$CAP1,prob=1-15/100),]
pro_spe <- rownames(pro_RDA_spe)
pro_spe

# 1999 - 2013 subset 
spe_sub <- spe.tib %>% filter(Site %in% c("caroldan", "isbasti", "iscaagua", "iscolon", "iscristo", "issobut")) %>% select(-c(Year)) %>% column_to_rownames("Site.Year") %>% select(-Site)

env_sub <- env.tib %>% filter(Site %in% c("caroldan", "isbasti", "iscaagua", "iscolon", "iscristo", "issobut")) %>% column_to_rownames("Site.Year")

sub_time <- capscale(log1p(spe_sub) ~ as.numeric(Year) + Condition(Site), env_sub,
                      add = TRUE, dist="bray")
summary(sub_time)
sub_R2adj <- RsquareAdj(sub_time)$adj.r.squared

sub_scores <- scores(sub_time)

# Top 15 % of species scores along the first axis (Year)
abs_sub <- as.data.frame(abs(sub_scores$species), keep.rownames=TRUE)
sub_RDA_spe <- abs_sub[abs_sub$CAP1 > quantile(abs_sub$CAP1,prob=1-15/100),]
sub_spe <- rownames(sub_RDA_spe)
sub_spe

sub_scores_dat <- as.data.frame(sub_scores$species)
sub_scores_dat[row.names(sub_scores_dat)[row.names(sub_scores_dat) %in% sub_spe],]
```