---
title: "Taxonomic & Functional Diversity"
author: "Julia Briand"
date: "1/31/2022"
output: html_document
---

This script was used to prepare data for Spatio-temporal patterns in coral reef composition and function across an altered environmental gradient. Contact: Julia Briand (julia.briand@mail.mcgill.ca)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(taxize)
library(adespatial)
library(rotl)
library(phytools)
library(ape)
library(Rphylopars)
library(apTreeshape)
library(geiger)
library(picante)
library(mice)
library(tidyverse)
library(lsr)
library(ggplot2)
library(betapart)
library(FD)
library(vegan)
```

## Trait data

```{r}
trait <- read.csv("Trait_Supp.csv", row.names = 1)
trait <- trait[,-c(1,4,5,7,8,12)] #remove categorical traits and notes
```

## Tree

```{r}
remove <- c("Isophyllastrea_rigida_MUS", "Madracis_asperula_AST", "Mycetophyllia_reesi_MUS", "Scolymia_lacera_MUS") #species not in tree
trait <- trait[!rownames(trait) %in% remove, ] #remove species with no tip

tree <- read.nexus(file = "doi_10.5061_dryad.178n3__v1/Huang&Roy_Molecular.tre")
keep <- keep.tip(tree, rownames(trait))
```

## Impute

```{r}
impute<-phylopars(trait,keep,phylo_correlated = FALSE) 
imput_trait <- impute$anc_recon[1:42,]

imput_trait <- rbind(imput_trait, trait[remove,])
imput_tib <- as_tibble(imput_trait, rownames = "tax")

# Add back the categorical trait columns
trait2 <- trait[, c(1,4,5,7,8,12)]
trait_tib <- as_tibble(trait2, rownames= "tax")
trait_tib <- trait_tib %>% na_if("")
full_trait <- trait_tib %>% left_join(imput_tib) 
full_trait <- full_trait %>% select(-c(tax))
full_trait <- full_trait %>% mutate(across(where(is.numeric), round, 3))
```

## Read in species and environemntal data

```{r}
env <- read.csv("Env_Supp.csv", row.names = 2) 
spe <- read.csv("Species.csv", row.names = 2)
```

## Subset only coral species

```{r}
spe_sub <- spe[,full_trait$species]
spe <- spe_sub
# mean species cover
spe.mean <- spe.mean[,full_trait$species]
```

## Transform species data

```{r}
spe.log <- log1p(spe)
spe.log.bray <- vegdist(spe.bray)
spe.mean.log <- log1p(spe.mean)
```

## Subset species data

```{r}
spe_sub <- spe[,full_trait$species]
full_trait <- full_trait %>% mutate(across(where(is.character), as.factor))
full_trait <- full_trait %>% column_to_rownames("species")
```

## Calculate matrices for Rao (diversity) function

```{r}
# Distance matrix based on a simplified phylogenetic classification
# Retrieve hierarchical classification from species list
splist <- c("Acropora cervicornis", "Acropora palmata",         
"Agaricia agaricites", "Agaricia fragilis",        
"Agaricia grahamae", "Agaricia humilis",         
"Agaricia lamarcki", "Agaricia tenuifolia",      
"Agaricia undata", "Colpophyllia natans",      
"Dichocoenia stokesii", "Diploria clivosa",         
"Diploria labyrinthiformis", "Diploria strigosa",        
"Eusmilia fastigiata", "Favia fragum",             
"Isophyllastrea rigida", "Isophyllia_sinuosa",       
"Helioseris cucullata", "Madracis asperula",        
"Madracis decactis", "Madracis auretenra",       
"Madracis pharensis", "Manicina areolata",        
"Meandrina meandrites", "Orbicella annularis",    
"Montastraea cavernosa", "Orbicella faveolata",    
"Orbicella franksi", "Mussa angulosa",           
"Mycetophyllia aliciae", "Mycetophyllia danaana",    
"Mycetophyllia lamarckiana", "Mycetophyllia reesi",      
"Porites astreoides", "Porites colonensis",       
"Porites divaricata", "Porites furcata",          
"Porites lobata", "Porites porites",          
"Scolymia cubensis", "Scolymia lacera",          
"Siderastrea radians", "Siderastrea siderea",      
"Solenastrea hyades", "Stephanocoenia intersepta")
get_gbifid(splist)
spcla <- classification(splist, db = "gbif")

# Compute the distance matrix and the phylogenetic tree
tr <- class2tree(spcla)

# Convert the tree to a cophenetic matrix
# constrained between 0 and 1
phylo.d <- cophenetic(tr$phylo) / 100

# Replace full species names by name codes
rownames(phylo.d) <- names(spe_sub)
colnames(phylo.d) <- names(spe_sub)

# Trait matrix
tra.new <- gowdis(full_trait)

# Test
spe.rao <- Rao(sample = log1p(t(spe_sub)), dfunc = trait.new, dphyl = phylo.d)
```

```{r}
sub_tib <- as_tibble(spe_sub, rownames = "Site.Year")
sub_tib <- sub_tib %>% separate(Site.Year, c("Site", "Year"))

sub_1999 <- sub_tib %>% filter(Year == 1999) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2000 <- sub_tib %>% filter(Year == 2000) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2001 <- sub_tib %>% filter(Year == 2001) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2002 <- sub_tib %>% filter(Year == 2002) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2003 <- sub_tib %>% filter(Year == 2003) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2004 <- sub_tib %>% filter(Year == 2004) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2005 <- sub_tib %>% filter(Year == 2005) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2006 <- sub_tib %>% filter(Year == 2006) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2007 <- sub_tib %>% filter(Year == 2007) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2008 <- sub_tib %>% filter(Year == 2008) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2009 <- sub_tib %>% filter(Year == 2009) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2010 <- sub_tib %>% filter(Year == 2010) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2011 <- sub_tib %>% filter(Year == 2011) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2012 <- sub_tib %>% filter(Year == 2012) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2013 <- sub_tib %>% filter(Year == 2013) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2008 <- sub_tib %>% filter(Year == 2008) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2009 <- sub_tib %>% filter(Year == 2009) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2010 <- sub_tib %>% filter(Year == 2010) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2011 <- sub_tib %>% filter(Year == 2011) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2012 <- sub_tib %>% filter(Year == 2012) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")
sub_2013 <- sub_tib %>% filter(Year == 2013) %>% 
  select(-c(Year)) %>% column_to_rownames("Site")



rao.1999 <- Rao(sample = log1p(t(sub_1999)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2000 <- Rao(sample = log1p(t(sub_2000)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2001 <- Rao(sample = log1p(t(sub_2001)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2002 <- Rao(sample = log1p(t(sub_2002)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2003 <- Rao(sample = log1p(t(sub_2003)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2004 <- Rao(sample = log1p(t(sub_2004)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2005 <- Rao(sample = log1p(t(sub_2005)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2006 <- Rao(sample = log1p(t(sub_2006)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2007 <- Rao(sample = log1p(t(sub_2007)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2008 <- Rao(sample = log1p(t(sub_2008)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2009 <- Rao(sample = log1p(t(sub_2009)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2010 <- Rao(sample = log1p(t(sub_2010)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2011 <- Rao(sample = log1p(t(sub_2011)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2012 <- Rao(sample = log1p(t(sub_2012)), 
                dfunc = tra.new, dphyl = phylo.d)
rao.2013 <- Rao(sample = log1p(t(sub_2013)), 
                dfunc = tra.new, dphyl = phylo.d)
```


```{r}
# Functional Diversity
alpha_1999 <- as_tibble_row(rao.1999$FD$Alpha)
alpha_2000 <- as_tibble_row(rao.2000$FD$Alpha)
alpha_2001 <- as_tibble_row(rao.2001$FD$Alpha)
alpha_2002 <- as_tibble_row(rao.2002$FD$Alpha)
alpha_2003 <- as_tibble_row(rao.2003$FD$Alpha)
alpha_2004 <- as_tibble_row(rao.2004$FD$Alpha)
alpha_2005 <- as_tibble_row(rao.2005$FD$Alpha)
alpha_2006 <- as_tibble_row(rao.2006$FD$Alpha)
alpha_2007 <- as_tibble_row(rao.2007$FD$Alpha)
alpha_fd2 <- bind_rows("1999"=alpha_1999,"2000"=alpha_2000,
                       "2001"=alpha_2001,"2002"=alpha_2002,
                       "2003"=alpha_2003,"2004"=alpha_2004,
                       "2005"=alpha_2005,"2006"=alpha_2006,
                       "2007"=alpha_2007,.id="Year")
alpha_fd2$Gallialm <- NA
alpha_fd2$bastinor <- NA
alpha_fd2$tobobe <- NA
alpha_fd2$zapoeste <- NA
alpha_fd2$Penivali <- NA
alpha_full <- full_join(alpha_fd2, alpha_fd)

longer_fd <- pivot_longer(alpha_full, c("zapoeste","isbasti","tobobe",
                                        "bastinor","caroldan", "Penivali", 
                                        "iscaagua", "iscolon", "iscristo", 
                                        "issobut", "Gallialm"),names_to = "Site")

longer_fd <- left_join(longer_fd, site_grps)

# Taxonomic Diversity 

alpha_2008 <- as_tibble_row(rao.2008$TD$Alpha)
alpha_2009 <- as_tibble_row(rao.2009$TD$Alpha)
alpha_2010 <- as_tibble_row(rao.2010$TD$Alpha)
alpha_2011 <- as_tibble_row(rao.2011$TD$Alpha)
alpha_2012 <- as_tibble_row(rao.2012$TD$Alpha)
alpha_2013 <- as_tibble_row(rao.2013$TD$Alpha)

alpha_2009$bastinor <- NA

alpha_td <- bind_rows("2008"=alpha_2008,"2009"=alpha_2009,
                      "2010"=alpha_2010,"2011"=alpha_2011,
                      "2012"=alpha_2012,"2013"=alpha_2013,
                      .id="Year")

alpha_1999 <- as_tibble_row(rao.1999$TD$Alpha)
alpha_2000 <- as_tibble_row(rao.2000$TD$Alpha)
alpha_2001 <- as_tibble_row(rao.2001$TD$Alpha)
alpha_2002 <- as_tibble_row(rao.2002$TD$Alpha)
alpha_2003 <- as_tibble_row(rao.2003$TD$Alpha)
alpha_2004 <- as_tibble_row(rao.2004$TD$Alpha)
alpha_2005 <- as_tibble_row(rao.2005$TD$Alpha)
alpha_2006 <- as_tibble_row(rao.2006$TD$Alpha)
alpha_2007 <- as_tibble_row(rao.2007$TD$Alpha)
alpha_td2 <- bind_rows("1999"=alpha_1999,"2000"=alpha_2000,
                       "2001"=alpha_2001,"2002"=alpha_2002,
                       "2003"=alpha_2003,"2004"=alpha_2004,
                       "2005"=alpha_2005,"2006"=alpha_2006,
                       "2007"=alpha_2007,.id="Year")
alpha_td2$Gallialm <- NA
alpha_td2$bastinor <- NA
alpha_td2$tobobe <- NA
alpha_td2$zapoeste <- NA
alpha_td2$Penivali <- NA
alpha_full2 <- full_join(alpha_td2, alpha_td)

longer_td <- pivot_longer(alpha_full2, c("zapoeste","isbasti","tobobe",
                                         "bastinor","caroldan", "Penivali", 
                                         "iscaagua", "iscolon", "iscristo", 
                                         "issobut", "Gallialm"),names_to = "Site")

longer_td <- left_join(longer_td, site_grps)

longer_td <- longer_td %>% mutate(Taxonomic = value) %>% select(-c(value))

longer_fd <- longer_fd %>% mutate(Functional = value) %>% select(-c(value))

longer_full <- full_join(longer_td, longer_fd)
# write.csv(longer_full, "TD_FD_Sites")
```

