---
title: "Temporal Models and Figures"
author: "Julia Briand"
date: "1/11/2022"
output: html_document
---

This script was used to prepare data for Spatio-temporal patterns in coral reef composition and function across an altered environmental gradient. Contact: Julia Briand (julia.briand@mail.mcgill.ca)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lme4)
library(nlme)
library(Matrix)
library(broom.mixed)
library(MuMIn)
library(lmerTest)
library(merTools)
```

## Read in Multivariate Regression Tree groups

```{r}
trait_grp <- read.csv("trait_mrt_grps.csv")
benth_grp <- read.csv("benth_mrt_grps.csv")
coral_grp <- read.csv("coral_mrt_grps.csv")

trait_grp <- trait_grp %>% 
  mutate(Site = mrt_sites, MRT = trait.mvpart.g) %>% 
  dplyr::select(-c(mrt_sites,trait.mvpart.g))

benth_grp <- benth_grp %>% 
  mutate(Site = X, MRT = benth.mvpart.g) %>% 
  dplyr::select(-c(X,benth.mvpart.g))

coral_grp <- coral_grp %>% 
  mutate(Site = X, MRT = coral.mvpart.g) %>% 
  dplyr::select(-c(X,coral.mvpart.g))
```

## Join groups with PCoA Scores

```{r}
# Corals
corals <- read.csv("Corals_PCoA.csv")

c_tib <- as_tibble(corals)
c_tib <- c_tib %>% mutate(Reef=recode(Group, `1` = "forereef", `2` = "lagoonal"))
c_tib$Reef <- as.factor(c_tib$Reef)
c_tib <- left_join(c_tib, coral_grp)
c_tib <- c_tib %>% mutate(MRT=recode(MRT, `1` = "node1", `2` = "node2", `3` = "node3", `4` = "node4"))
c_tib$MRT <- as.factor(c_tib$MRT)

# Benthic
benthic <- read.csv("Taxa_PCoA.csv")

b_tib <- as_tibble(benthic)
b_tib <- b_tib %>% mutate(Reef=recode(Group, `1` = "forereef", `2` = "lagoonal"))
b_tib$Reef <- as.factor(b_tib$Reef)
b_tib <- left_join(b_tib, benth_grp)
b_tib <- b_tib %>% mutate(MRT=recode(MRT, `1` = "node1", `2` = "node2", `3` = "node3"))
b_tib$MRT <- as.factor(b_tib$MRT)

# Traits
traits <- read.csv("Trait_MFA.csv")

t_tib <- as_tibble(traits)
t_tib <- t_tib %>% mutate(Reef=recode(Alternate, `2` = "forereef", `1` = "lagoonal"))
t_tib$Reef <- as.factor(t_tib$Reef)
t_tib <- left_join(t_tib, trait_grp)
t_tib <- t_tib %>% mutate(MRT=recode(MRT, `1` = "node1", `2` = "node2", `3` = "node3", `4` = "node4"))
t_tib$MRT <- as.factor(t_tib$MRT)
```

## Run PCoA1 temporal models

# Coral

```{r}
# Random Intercept
p1_c1 <- lme4::lmer(MDS1 ~ Year*MRT + (1|MRT/Site), data = c_tib)
summary(p1_c1)
summary(as_lmerModLmerTest(p1_c1))

# No Interaction Term
p1_c2 <- lme4::lmer(MDS1 ~ Year+MRT + (1|MRT/Site), data = c_tib)
summary(p1_c2)
summary(as_lmerModLmerTest(p1_c2))

# Anova
anova(p1_c1, p1_c2)

#lme
p1_lmec1 <- lme(fixed = MDS1~Year*MRT, random = ~ 1|MRT/Site,data=c_tib,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(c_tib, nesting(MRT, Site), Year) 
pred<-predict(p1_lmec1, newdata, level=0, se.fit=T) 
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, c_tib)

p1_c <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, MDS1, color = Reef, shape = MRT),c_tib, size = 2.5)+geom_path(aes(Year, MDS1, group= Site, color = Reef),c_tib[!is.na(c_tib$MDS1),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + scale_fill_manual(values=c("#FEB24C", "#FF4500", "#FF4500", "#FF4500")) + geom_path() + 
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + scale_shape_manual(values=c(15, 16, 17, 18)) +
  labs(x = "Year", y = "PCoA 1")+facet_wrap(~ MRT,scales='free') + ylim(-2.0, 2.0)+theme(legend.position = "none")
  
ggsave("pcoa1_coral_mrt.pdf", height=7, width=7.17)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Benthic

```{r}
# Random Intercept
p1_b1 <- lme4::lmer(MDS1 ~ Year*MRT + (1|MRT/Site), data = b_tib)
summary(p1_b1)
summary(as_lmerModLmerTest(p1_b1))

# No Interaction Term
p1_b2 <- lme4::lmer(MDS1 ~ Year+MRT + (1|MRT/Site), data = b_tib)
summary(p1_b2)
summary(as_lmerModLmerTest(p1_b2))

# Anova
anova(p1_b1, p1_b2)

#lme
p1_lmeb1 <- lme(fixed = MDS1~Year*MRT, random = ~ 1|MRT/Site,data=b_tib,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(b_tib, nesting(MRT, Site), Year) 
pred<-predict(p1_lmeb1, newdata, level=0, se.fit=T) 
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, b_tib)

p1_b <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, MDS1, color = Reef, shape = MRT),b_tib, size = 2.5)+geom_path(aes(Year, MDS1, group= Site, color = Reef),b_tib[!is.na(b_tib$MDS1),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + scale_fill_manual(values=c("#FEB24C", "#FF4500", "#FF4500", "#FF4500")) + geom_path() + 
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + scale_shape_manual(values=c(15, 16, 17, 18)) + geom_path() + labs(x = "Year", y = "PCoA 1")+
  facet_wrap(~ MRT,scales='free') + ylim(-2.0, 2.0) +theme(legend.position = "none")

ggsave("pcoa1_benth_mrt.pdf", height=4.5, width=14)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Traits

```{r}
# Random Intercept
p1_t1 <- lme4::lmer(Dim.1 ~ Year*MRT + (1|MRT/Site), data = t_tib)
summary(p1_t1)
summary(as_lmerModLmerTest(p1_t1))

# No Interaction Term
p1_t2 <- lme4::lmer(Dim.1 ~ Year+MRT + (1|MRT/Site), data = t_tib)
summary(p1_t2)
summary(as_lmerModLmerTest(p1_t2))

# Anova
anova(p1_t1, p1_t2)

#lme
p1_lmet1 <- lme(fixed = Dim.1~Year*MRT, random = ~ 1|MRT/Site,data=t_tib,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(t_tib, nesting(MRT, Site), Year) 
pred<-predict(p1_lmet1, newdata, level=0, se.fit=T) 
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, t_tib)

p1_t <- ggplot(fitted_pred, aes(Year, fit)) + geom_path(size = 0.8)+geom_point(aes(Year, Dim.1, color = Reef, shape = MRT),t_tib, size = 2.5)+geom_path(aes(Year, Dim.1, group= Site, color = Reef),t_tib[!is.na(t_tib$Dim.1),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + 
  scale_fill_manual(values=c("#FEB24C", "#FF4500", "#FF4500", "#FF4500")) + geom_path() +
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + 
  scale_shape_manual(values=c(15, 16, 17, 18)) + 
  labs(x = "Year", y = "MFA 1") +
  facet_wrap(~ MRT,scales='free') + 
  ylim(-6.0, 6.0) +theme(legend.position = "none")

ggsave("pcoa1_trait_mrt.pdf", height=7, width=7.17)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

## PCOA2 Models

# Coral

```{r}
# Corals

# Random Intercept
p2_c1 <- lme4::lmer(MDS2 ~ Year*MRT + (1|MRT/Site), data = c_tib)
summary(p2_c1)
summary(as_lmerModLmerTest(p2_c1))

# No Interaction Term
p2_c2 <- lme4::lmer(MDS2 ~ Year+MRT + (1|MRT/Site), data = c_tib)
summary(p2_c2)
summary(as_lmerModLmerTest(p2_c2))

# Anova
anova(p2_c1, p2_c2)

#lme
p2_lmec1 <- lme(fixed = MDS2~Year*MRT, random = ~ 1|MRT/Site,data=c_tib,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(c_tib, nesting(MRT, Site), Year) 
pred<-predict(p2_lmec1, newdata, level=0, se.fit=T) 
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, c_tib)

p2_c <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, MDS2, color = Reef, shape = MRT),c_tib, size = 2.5)+geom_path(aes(Year, MDS2, group= Site, color = Reef),c_tib[!is.na(c_tib$MDS2),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + labs(x = "Year", y = "PCoA 2")+ scale_fill_manual(values=c("#FEB24C", "#FF4500", "#FF4500", "#FF4500")) + geom_path() + 
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + scale_shape_manual(values=c(15, 16, 17, 18)) +theme(legend.position = "none") + facet_wrap(~ MRT,scales='free') + ylim(-2.5, 2.5) 

ggsave("pcoa2_coral_mrt.pdf", height=8.5, width=10.75)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Benthic

```{r}
# Random Intercept
p2_b1 <- lme4::lmer(MDS2 ~ Year*MRT + (1|MRT/Site), data = b_tib)
summary(p2_b1)
summary(as_lmerModLmerTest(p2_b1))

# No Interaction Term
p2_b2 <- lme4::lmer(MDS2 ~ Year+MRT + (1|MRT/Site), data = b_tib)
summary(p2_b2)
summary(as_lmerModLmerTest(p2_b2))

# Anova
anova(p2_b1, p2_b2)

#lme
p2_lmeb1 <- lme(fixed = MDS2~Year*MRT, random = ~ 1|MRT/Site,data=b_tib,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(b_tib, nesting(MRT, Site), Year) 
pred<-predict(p2_lmeb1, newdata, level=0, se.fit=T) 
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, b_tib)

p2_b <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, MDS2, color = Reef, shape = MRT),b_tib, size = 2.5)+geom_path(aes(Year, MDS2, group= Site, color = Reef),b_tib[!is.na(b_tib$MDS2),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5) +
  ggthemes::theme_base() + 
  labs(x = "Year", y = "PCoA 2") + 
  scale_fill_manual(values=c("#FEB24C", "#FF4500", "#FF4500", "#FF4500")) + 
  geom_path() + 
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + 
  scale_shape_manual(values=c(15, 16, 17, 18)) +
  theme(legend.position = "none")+
  facet_wrap(~ MRT,scales='free') + ylim(-2.0, 2.0) 

ggsave("pcoa2_benth_mrt.pdf", height=4.5, width=15)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Traits

```{r}
# Random Intercept
p2_t1 <- lme4::lmer(Dim.2 ~ Year*MRT + (1|MRT/Site), data = t_tib)
summary(p2_t1)
summary(as_lmerModLmerTest(p2_t1))

# No Interaction Term
p2_t2 <- lme4::lmer(Dim.2 ~ Year+MRT + (1|MRT/Site), data = t_tib)
summary(p2_t2)
summary(as_lmerModLmerTest(p2_t2))

# Anova
anova(p2_t1, p2_t2)

#lme
p2_lmet1 <- lme(fixed = Dim.2~Year*MRT, random = ~ 1|MRT/Site,data=t_tib,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(t_tib, nesting(MRT, Site), Year) 
pred<-predict(p2_lmet1, newdata, level=0, se.fit=T) 
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, t_tib)

p2_t <- ggplot(fitted_pred, aes(Year, fit)) + geom_path(size = 0.8)+geom_point(aes(Year, Dim.2, color = Reef, shape = MRT),t_tib, size = 2.5)+geom_path(aes(Year, Dim.2, group= Site, color = Reef),t_tib[!is.na(t_tib$Dim.2),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + 
  labs(x = "Year", y = "MFA 2")+
  scale_fill_manual(values=c("#FEB24C", "#FF4500", "#FF4500", "#FF4500")) + 
  geom_path() + 
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + 
  scale_shape_manual(values=c(15, 16, 17, 18)) +
  theme(legend.position = "none")+
  facet_wrap(~ MRT,scales='free') + ylim(-6.0, 6.0) 

ggsave("pcoa2_trait_mrt.pdf", height=4.5, width=15)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

## Diversity Models

# Taxonomic Diversity

```{r}
# Read in diversity for each site every year
Div <- read.table("TD_FD_Sites",sep=",",header=T)
Div <- Div[,-1]
div_tib <- as_tibble(Div)
div_tib <- div_tib %>% mutate(Reef=recode(Alternate, `1` = "lagoonal", `2` = "forereef"))
div_tib$Reef <- as.factor(div_tib$Reef)

b_div <- left_join(div_tib, benth_grp)
b_div <- b_div %>% mutate(MRT=recode(MRT, `1` = "node1", `2` = "node2", `3` = "node3"))
b_div$MRT <- as.factor(b_div$MRT)

t_div <- left_join(div_tib, trait_grp)
t_div <- t_div %>% mutate(MRT=recode(MRT, `1` = "node1", `2` = "node2", `3` = "node3", `4` = "node4"))
t_div$MRT <- as.factor(t_div$MRT)

c_div <- left_join(div_tib, coral_grp)
c_div <- c_div %>% mutate(MRT=recode(MRT, `1` = "node1", `2` = "node2", `3` = "node3", `4` = "node4"))
c_div$MRT <- as.factor(c_div$MRT)
```

# Corals

```{r}
# Random Intercept
d_c1 <- lme4::lmer(Taxonomic ~ Year*MRT + (1|MRT/Site), data = c_div)
summary(d_c1)
summary(as_lmerModLmerTest(d_c1))

# No Interaction Term
d_c2 <- lme4::lmer(Taxonomic ~ Year+MRT + (1|MRT/Site), data = c_div)
summary(d_c2)
summary(as_lmerModLmerTest(d_c2))

# Anova
anova(d_c1, d_c2)

#lme
d_lmec1 <- lme(fixed = Taxonomic~Year*MRT, random = ~ 1|MRT/Site,data=c_div,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(c_div, nesting(MRT, Site), Year) 
pred<-predict(d_lmec1, newdata, level=0, se.fit=T) 
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, c_div)

d_t <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, Taxonomic, color = Reef, shape = MRT),c_div, size = 2.5)+geom_path(aes(Year, Taxonomic, group= Site, color = Reef),c_div[!is.na(c_div$Taxonomic),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = Reef), fitted_pred, alpha = .5) +
  ggthemes::theme_base() +
  scale_fill_manual(values=c("#FEB24C", "#FF4500")) + 
  geom_path() + labs(x = "Year", y = "Taxonomic Diversity") +
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + 
  scale_shape_manual(values=c(15, 16, 17, 18)) +
  theme(legend.position = "none") + 
  facet_wrap(~ MRT,scales='free') + ylim(0.7, 1.1) 

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Benthic

```{r}
# Random Intercept
d_b1 <- lme4::lmer(Taxonomic ~ Year*MRT + (1|MRT/Site), data = b_div)
summary(d_b1)
summary(as_lmerModLmerTest(d_b1))

# No Interaction Term
d_b2 <- lme4::lmer(Taxonomic ~ Year+MRT + (1|MRT/Site), data = b_div)
summary(d_b2)
summary(as_lmerModLmerTest(d_b2))

# Anova
anova(d_b1, d_b2)

#lme
d_lmeb1 <- lme(fixed = Taxonomic~Year*MRT, random = ~ 1|MRT/Site,data=b_div,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(b_div, nesting(MRT, Site), Year) 
pred<-predict(d_lmeb1, newdata, level=0, se.fit=T)
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, b_div)

d_b <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, Taxonomic, color = Reef, shape = MRT),b_div, size = 2.5)+geom_path(aes(Year, Taxonomic, group= Site, color = Reef),b_div[!is.na(b_div$Taxonomic),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = Reef), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + 
  scale_fill_manual(values=c("#FF8C00", "#FF4500")) + 
  geom_path() + labs(x = "Year", y = "Taxonomic Diversity")+
  scale_color_manual(values=c("#FF8C00", "#FF4500")) + 
  facet_wrap(~ MRT, scales='free') + ylim(0.7, 1.1) 

ggsave("tax_benth_mrt.pdf", height=4.5, width=15)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Traits

```{r}
# Random Intercept
d_t1 <- lme4::lmer(Taxonomic ~ Year*MRT + (1|MRT/Site), data = t_div)
summary(d_t1)
summary(as_lmerModLmerTest(d_t1))

# No Interaction Term
d_t2 <- lme4::lmer(Taxonomic ~ Year+MRT + (1|MRT/Site), data = t_div)
summary(d_t2)
summary(as_lmerModLmerTest(d_t2))

# Anova
anova(d_t1, d_t2)

#lme
d_lmet1 <- lme(fixed = Taxonomic~Year*MRT, random = ~ 1|MRT/Site,data=t_div,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(t_div, nesting(MRT, Site), Year) 
pred<-predict(d_lmet1, newdata, level=0, se.fit=T)
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, t_div)

d_t <- ggplot(fitted_pred, aes(Year, fit)) + geom_path(size = 0.8)+geom_point(aes(Year, Taxonomic, color = Reef, shape = MRT),t_div, size = 2.5)+geom_path(aes(Year, Taxonomic, group= Site, color = Reef),t_div[!is.na(t_div$Taxonomic),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5) +
  ggthemes::theme_base() + 
  scale_fill_manual(values=c("#FF8C00", "#FF4500", "#FF4500")) + 
  geom_path() + labs(x = "Year", y = "Taxonomic Diversity") +
  scale_color_manual(values=c("#FF8C00", "#FF4500")) + 
  facet_wrap(~ MRT, scales='free') + ylim(0.7, 1.1)

ggsave("tax_trait_mrt.pdf", height=4.5, width=15)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Functional Diversity

# Corals

```{r}
# Random Intercept
f_c1 <- lme4::lmer(Functional ~ Year*MRT + (1|MRT/Site), data = c_div)
summary(f_c1)
summary(as_lmerModLmerTest(f_c1))

# No Interaction Term
f_c2 <- lme4::lmer(Functional ~ Year+MRT + (1|MRT/Site), data = c_div)
summary(f_c2)
summary(as_lmerModLmerTest(f_c2))

# Anova
anova(f_c1, f_c2)

#lme
f_lmec1 <- lme(fixed = Functional~Year*MRT, random = ~ 1|MRT/Site,data=c_div,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(c_div, nesting(MRT, Site), Year) 
pred<-predict(f_lmec1, newdata, level=0, se.fit=T)
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, c_div)

f_c <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, Functional, color = Reef, shape = MRT),c_div, size = 2.5)+geom_path(aes(Year, Functional, group= Site, color = Reef),c_div[!is.na(c_div$Functional),], size = 0.7) +
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = Reef), fitted_pred, alpha = .5) +
  ggthemes::theme_base() + 
  scale_fill_manual(values=c("#FF8C00", "#FF4500")) + 
  geom_path() + labs(x = "Year", y = "Functional Diversity") +
  scale_color_manual(values=c("#FF8C00", "#FF4500")) + 
  facet_wrap(~ MRT, scales='free') + ylim(0.1, 0.4)

ggsave("fun_coral_mrt.pdf", height=8.5, width=10.75)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Benthic

```{r}
# Random Intercept
f_b1 <- lme4::lmer(Functional ~ Year*MRT + (1|MRT/Site), data = b_div)
summary(f_b1)
summary(as_lmerModLmerTest(f_b1))

# No Interaction Term
f_b2 <- lme4::lmer(Functional ~ Year+MRT + (1|MRT/Site), data = b_div)
summary(b2)
summary(as_lmerModLmerTest(b2))

# Anova
anova(f_b1, f_b2)

#lme
f_lmeb1 <- lme(fixed = Functional~Year*MRT, random = ~ 1|MRT/Site,data=b_div,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(b_div, nesting(MRT, Site), Year) 
pred<-predict(f_lmeb1, newdata, level=0, se.fit=T)
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, b_div)

f_b <- ggplot(fitted_pred, aes(Year, fit, group = MRT)) + geom_path(size = 0.8)+geom_point(aes(Year, Functional, color = Reef, shape = MRT),b_div, size = 2.5)+geom_path(aes(Year, Functional, group= Site, color = Reef),b_div[!is.na(b_div$Functional),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = Reef), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + 
  scale_fill_manual(values=c("#FF8C00", "#FF4500")) + 
  geom_path() + labs(x = "Year", y = "Functional Diversity")+
  scale_color_manual(values=c("#FF8C00", "#FF4500")) + 
  facet_wrap(~ MRT, scales='free') + ylim(0.1, 0.4) 

ggsave("fun_benth_mrt.pdf", height=4.5, width=15)

#rm(newdata, pred, fitted_pred, fitted_pred)
```

# Traits

```{r}
# Random Intercept
f_t1 <- lme4::lmer(Functional ~ Year*MRT + (1|MRT/Site), data = t_div)
summary(f_t1)
summary(as_lmerModLmerTest(f_t1))

# No Interaction Term
f_t2 <- lme4::lmer(Functional ~ Year+MRT + (1|MRT/Site), data = t_div)
summary(f_t2)
summary(as_lmerModLmerTest(f_t2))

# Anova
anova(f_t1, f_t2)

#lme
f_lmet1 <- lme(fixed = Functional~Year*MRT, random = ~ 1|MRT/Site,data=t_div,na.action=na.omit)

# taxonomic c1
newdata <- tidyr::expand(t_div, nesting(MRT, Site), Year) 
pred<-predict(f_lmet1, newdata, level=0, se.fit=T)
fitted_pred<-newdata %>%
  mutate(fit=pred$fit) %>%
  mutate(fit_se=pred$se.fit)

fitted_pred <- left_join(fitted_pred, t_div)

f_p <- ggplot(fitted_pred, aes(Year, fit)) + geom_path(size = 0.8)+geom_point(aes(Year, Functional, color = Reef, shape = MRT),t_div, size = 2.5)+geom_path(aes(Year, Functional, group= Site, color = Reef),t_div[!is.na(t_div$Functional),], size = 0.7)+
  geom_ribbon( aes(ymin = fit-1.96*fit_se, ymax = fit+1.96*fit_se, fill = MRT), fitted_pred, alpha = .5)+
  ggthemes::theme_base() + scale_fill_manual(values=c("#FEB24C", "#FF4500", "#FF4500", "#FF4500")) + geom_path() + 
  scale_color_manual(values=c("#FEB24C", "#FF4500")) + 
  scale_shape_manual(values=c(15, 16, 17, 18)) +
  theme(legend.position = "none") + 
  labs(x = "Year", y = "Functional Diversity") +
  facet_wrap(~ MRT, scales='free') + ylim(0.1, 0.4) 

ggsave("fun_trait_mrt.pdf", height=4.5, width=15)

#rm(newdata, pred, fitted_pred, fitted_pred)
```


