---
title: "Trait Imputation, Clustering & Ordination"
author: "Julia Briand"
date: "1/27/2022"
output: html_document
---

This script was used to prepare data for Spatio-temporal patterns in coral reef composition and function across an altered environmental gradient. Contact: Julia Briand (julia.briand@mail.mcgill.ca)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rotl)
library(phytools)
library(ape)
library(Rphylopars)
library(apTreeshape)
library(geiger)
library(picante)
library(mice)
library(tidyverse)
library(lsr)
library(ggplot2)
library(betapart)
library(FD)
library(vegan)
library(FactoMineR)
library(ggrepel)
```

## Trait data

```{r}
trait <- read.csv("Trait_Supp.csv", row.names = 1)
trait <- trait[,-c(1,4,5,7,8,12)] #remove categorical traits and notes
```

## Tree

```{r}
remove <- c("Isophyllastrea_rigida_MUS", "Madracis_asperula_AST", "Mycetophyllia_reesi_MUS", "Scolymia_lacera_MUS") #species not in tree
trait <- trait[!rownames(trait) %in% remove, ] #remove species with no tip

tree <- read.nexus(file = "doi_10.5061_dryad.178n3__v1/Huang&Roy_Molecular.tre")
keep <- keep.tip(tree, rownames(trait))
```

## Impute

```{r}
impute<-phylopars(trait,keep,phylo_correlated = FALSE) 
imput_trait <- impute$anc_recon[1:42,]

imput_trait <- rbind(imput_trait, trait[remove,])
imput_tib <- as_tibble(imput_trait, rownames = "tax")

# Add back the categorical trait columns
trait2 <- trait[, c(1,4,5,7,8,12)]
trait_tib <- as_tibble(trait2, rownames= "tax")
trait_tib <- trait_tib %>% na_if("")
full_trait <- trait_tib %>% left_join(imput_tib) 
full_trait <- full_trait %>% select(-c(tax))
full_trait <- full_trait %>% mutate(across(where(is.numeric), round, 3))
```

## Read in species and environemntal data

```{r}
env <- read.csv("Env_Supp.csv", row.names = 2) 
spe <- read.csv("Species.csv", row.names = 2)
```

## Subset only coral species

```{r}
spe_sub <- spe[,full_trait$species]
spe <- spe_sub
# mean species cover
spe.mean <- spe.mean[,full_trait$species]
```

## Transform species data

```{r}
spe.log <- log1p(spe)
spe.log.bray <- vegdist(spe.bray)
spe.mean.log <- log1p(spe.mean)
```

## Standardize Environmental Data

```{r}
env.z <- decostand(env, method="standardize")
# the data are now centered (means~0)
apply(env.z, 2, mean) 
# the data are now scaled (standard deviations=1)
apply(env.z, 2, sd)
```

## Community Weighted Means

```{r}
full_trait <- full_trait %>% mutate(across(where(is.character), as.factor))
full_trait <- full_trait %>% column_to_rownames("species")
cwm <- functcomp(full_trait, as.matrix(spe_sub), CWM.type = "all")
scale_cwm <- scale(cwm)
```

## K means clustering

```{r}
cwm_tib <- as_tibble(cwm, rownames = "Site.Year")
cwm_tib <- cwm_tib %>% separate(Site.Year, c("Site", "Year"), remove = FALSE)
cwm_tib <- cwm_tib %>% mutate_at(vars(colnames(cwm_tib[4:ncol(cwm_tib)])), as.numeric)

sapply(cwm_tib, is.numeric) # check

bastinor <- cwm_tib %>% 
filter(Site ==  "bastinor") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
caroldan <- cwm_tib %>% 
filter(Site == "caroldan") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
Gallialm <- cwm_tib %>% 
filter(Site == "Gallialm") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
isbasti <- cwm_tib %>% filter(Site == "isbasti") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
iscaagua <- cwm_tib %>% filter(Site == "iscaagua") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
iscolon <- cwm_tib %>% filter(Site == "iscolon") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
iscristo <- cwm_tib %>% filter(Site == "iscristo") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
issobut <- cwm_tib %>% filter(Site == "issobut") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
Penivali <- cwm_tib %>% filter(Site == "Penivali") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
tobobe <- cwm_tib %>% filter(Site == "tobobe") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
zapoeste <- cwm_tib %>% filter(Site == "zapoeste") %>% 
summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))

# Combine site tibbles:
site.mean <- bind_rows(zapoeste = zapoeste,tobobe = tobobe,issobut = issobut,iscristo = iscristo,iscolon = iscolon,iscaagua = iscaagua,isbasti = isbasti,Gallialm = Gallialm, caroldan = caroldan, bastinor = bastinor,Penivali = Penivali, .id = "Site")

#' Reorder:
site.mean <- site.mean %>% column_to_rownames(var = "Site")

cwm_mean_stand <- decostand(site.mean, "range")
cwm_mean_scale <- scale(site.mean)
spe.kmeans <- kmeans(cwm_mean_scale, centers = 2, nstart = 100000)
site_grps <- as.data.frame(unlist(spe.kmeans$cluster))
names(site_grps) <- (c("Group"))
site_grps <- rownames_to_column(site_grps, var = "Site")
site_grps
```

## Multivariate Regression Tree

```{r}
trait.mean.mrt <- mvpart(as.matrix(cwm_mean_scale) ~SST.Mean + SPM + CHL + Fetch + Depth,env.mean,
                         legend=FALSE, margin=0.01, cp=0, xv="pick",
                         xval=nrow(cwm_mean_scale), xvmult=100, which=4,size =4)

summary(trait.mean.mrt)
printcp(trait.mean.mrt)

#MRT groups
trait.mean.MRT <- MRT(trait.mean.mrt)
trait.mvpart.g <- factor(trait.mean.mrt$where)
levels(trait.mvpart.g) <- 1:length(levels(trait.mvpart.g))
trait.mvpart.g
mrt_sites <- rownames(cwm_mean_scale)
mrt_dat <- as.data.frame(mrt_sites)
mrt_dat <- cbind(mrt_dat, trait.mvpart.g)
```

## MFA

```{r}
grp <- c(7, 2, 2, 3, 3, 1, 1, 1, 1, 1)
# Compute the MFA
mfa <- MFA(scale_cwm,
              group = grp,
              type = c("c", "c", "c", "c", "c", "c", "c", "c", "c", "c"),
              ncp = 50,
              name.group = c("growth_form", "larval_dev", "sex_syst", "water_clar", 
                             "wave_exp", "max_colony", "growth_rate", 
                             "corallite_max", "depth_up", "depth_range"))

# Top 30 % of trait scores along the first axis
mfa$global.pca$var$coord # coordinates of traits
abs_trait <- as.data.frame(abs(mfa$global.pca$var$coord), keep.rownames=TRUE)
Dim.1_trait <- abs_trait[abs_trait$Dim.1 > quantile(abs_trait$Dim.1,prob=1-30/100),]

# Top 30 % of trait scores along the second axis
Dim.2_trait <- abs_trait[abs_trait$Dim.2 > quantile(abs_trait$Dim.2,prob=1-30/100),]
MFA_trait <- rownames(Dim.1_trait)

mfa_dat <- as.data.frame(mfa$global.pca$var$coord[MFA_trait,])
ggplot(mfa_dat, aes(Dim.1, Dim.2)) + 
  geom_segment(aes(x = rep(0,7),  y = rep(0,7), 
                   xend = Dim.1, yend = Dim.2), 
               data = mfa_dat, arrow=arrow(length = unit(0.3,"cm"))) +
  geom_text_repel(aes(label = rownames(mfa_dat)), 
                  data = mfa_dat, box.padding = 0.2, 
                  nudge_y = -0.01, nudge_x = -0.04) + 
  ggthemes::theme_base() +
  xlim(-6,6) + ylim(-6,6)

ggsave("trait_scores.pdf", height=5, width=6)

mfa$global.pca$ind$coord #coords for the sites & years

mfa_site <- as.data.frame(mfa$global.pca$ind$coord)
mfa_tib <- as_tibble(mfa_site, rownames = "sites")
mfa_tib <- mfa_tib %>% separate(sites, c("Site", "Year"), remove = FALSE)

mfa_grp <- mfa_tib %>% left_join(site_grps)
ggplot(mfa_grp, aes(Dim.1, Dim.2, color = as.character(Group), group = Group)) +
  geom_point(size = 2.5) +
  ggthemes::theme_base()+
scale_color_manual(values=c("#FF99CC", "#56B4E9")) +
  xlim(-5.5,5.5) + ylim(-6,6) +
  labs(x = "MFA 1",
       y = "MFA 2")

```

## RDA

```{r}
sst_rda <- rda(scale(cwm) ~ SST.Mean + SPM + CHL + Fetch + Depth,
         env.z)

summary(sst_rda, display=NULL)

# R2
sst_R2adj <- RsquareAdj(sst_rda)$adj.r.squared #0.3186435

# Model Selection
ordiR2step(rda(scale(cwm) ~ 1, data=env.z), scope= formula(sst_rda), direction= "forward", R2scope=TRUE, pstep=1000)

# Significance of Axes
anova.cca(sst_rda, step=1000) 
anova.cca(sst_rda, step=1000, by="axis")

# Top 30 % of trait scores along the first axis
abs_sst <- as.data.frame(abs(sst_scores$species), keep.rownames=TRUE)
sst_RDA1 <- abs_sst[abs_sst$RDA1 > quantile(abs_sst$RDA1,prob=1-30/100),]
sst_RDA_traits <- rownames(sst_RDA1)
sst_RDA_traits
```

## Temporal (Partial) RDA

```{r}
env.tib <- as_tibble(env, rownames = "Site.Year")
env.tib <- env.tib %>% separate(Site.Year, c("Site", "Year"), remove = FALSE)
env.tab <- env.tib %>% column_to_rownames(var = "Site.Year") 
env.tib <- env.tib %>% mutate(Year = as.numeric(Year))

# Exposed
cwm_exp <- cwm_tib %>% filter(Year >= 2008) %>% select(-c(Year)) %>%  
  filter(Site %in% c("zapoeste", "isbasti", "tobobe", "bastinor")) %>% column_to_rownames("Site.Year") %>% select(-Site)

# Protected
cwm_pro <- cwm_tib %>% filter(Year >= 2008) %>% select(-c(Year)) %>%  
  filter(Site %in% c("caroldan", "Penivali", "iscaagua", "iscolon", "iscristo", "issobut", "Gallialm")) %>% column_to_rownames("Site.Year") %>% select(-Site)

# Subset Env Data

# Exp Env
env_exp <- env.tib %>% filter(Year >= 2008) %>% 
  filter(Site %in% c("zapoeste", "isbasti", "tobobe", "bastinor")) %>% column_to_rownames("Site.Year")

# Pro Env
env_pro <- env.tib %>% filter(Year >= 2008) %>%
  filter(Site %in% c("caroldan", "Penivali", "iscaagua", "iscolon", "iscristo", "issobut", "Gallialm")) %>% column_to_rownames("Site.Year")

# Partial dbRDA Exposed
exp_time <- rda(scale(cwm_exp) ~ as.numeric(Year) + Condition(Site), env_exp)
summary(exp_time)
exp_R2adj <- RsquareAdj(exp_time)$adj.r.squared #0.003968873

# Partial DB RDA scores
exp_scores <- scores(exp_time)

# Top 30 % of trait scores along the first axis (Year)
abs_exp <- as.data.frame(abs(exp_scores$species), keep.rownames=TRUE)
exp_RDA_spe <- abs_exp[abs_exp$RDA1 > quantile(abs_exp$RDA1,prob=1-30/100),]
exp_spe <- rownames(exp_RDA_spe)
exp_spe

exp_scores$species[rownames(exp_RDA_spe),]

# Partial dbRDA Protected
pro_time <- rda(scale(cwm_pro) ~ as.numeric(Year) + Condition(Site), env_pro)
summary(pro_time)
pro_R2adj <- RsquareAdj(pro_time)$adj.r.squared #0.04009446

# Partial DB RDA scores
pro_scores <- scores(pro_time)

# Top 30 % of species scores along the first axis (Year)
abs_pro <- as.data.frame(abs(pro_scores$species), keep.rownames=TRUE)
pro_RDA_spe <- abs_pro[abs_pro$RDA1 > quantile(abs_pro$RDA1,prob=1-30/100),]
pro_spe <- rownames(pro_RDA_spe)
pro_spe

pro_scores$species[rownames(pro_RDA_spe),]

# 1999 - 2013 subset 
cwm_sub <- cwm_tib %>% filter(Site %in% 
                                c("caroldan", "isbasti", "iscaagua", "iscolon", "iscristo", "issobut")) %>% 
  select(-c(Year)) %>% 
  column_to_rownames("Site.Year") %>% 
  select(-Site)

env_sub <- env.tib %>% filter(Site %in% 
                                c("caroldan", "isbasti", "iscaagua", "iscolon", "iscristo", "issobut")) %>% 
  column_to_rownames("Site.Year")

sub_time <- capscale(log1p(cwm_sub) ~ as.numeric(Year) + Condition(Site), env_sub,
                     add = TRUE, dist="bray")
summary(sub_time)
sub_R2adj <- RsquareAdj(sub_time)$adj.r.squared

sub_scores <- scores(sub_time)

# Top 30 % of species scores along the first axis (Year)
abs_sub <- as.data.frame(abs(sub_scores$species), keep.rownames=TRUE)
sub_RDA_spe <- abs_sub[abs_sub$CAP1 > quantile(abs_sub$CAP1,prob=1-30/100),]
sub_spe <- rownames(sub_RDA_spe)
sub_spe

sub_scores_dat <- as.data.frame(sub_scores$species)
sub_scores_dat[row.names(sub_scores_dat)[row.names(sub_scores_dat) %in% sub_spe],]
```

